---
title: "project final report Appendix"
author: "Ailing Liu"
date: "2025-10-25"
output: pdf_document
---

```{r,warning=FALSE,message=FALSE}
library(tidyverse)
library(knitr)
library(purrr)
library(ggplot2)
library(corrplot)
library(forecast)
library(astsa)
library(vars)
library(dplyr)
library(conflicted)



conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::filter)


df2 <- read.csv("C:/Users/20871/Desktop/utsc/fourth year/stad57/project/Project_Group3/2023 monthly data.csv")

#variable selecting and data cleaning
df1 <- df2 %>% filter(Country %in% c("Canada","United States"), 
                      BALANCE %in% c("COMBFUEL","NUCLEAR","HYDRO"))
df1 <- df1 %>% select(Country,BALANCE,TIME,Value)
```
# 1. Introduction

# 2. Dataset
```{r,eval=FALSE}
#statistic summary(we calculate population standard deviation)
df3 <- df1 %>% group_by(Country,BALANCE) %>% 
  summarise(Mean = mean(Value, na.rm = TRUE),
            SD=sqrt(mean((Value - mean(Value))^2)),
            Min = min(Value),
            Max = max(Value))

kable(df3)
```

\pagebreak

\begin{center}
\textbf{Table A1: Statistic Summary of Time Series}
\end{center}
```{r,echo=FALSE,message=FALSE}
#statistic summary(we calculate population standard deviation)
df3 <- df1 %>% group_by(Country,BALANCE) %>% 
  summarise(Mean = mean(Value, na.rm = TRUE),
            SD=sqrt(mean((Value - mean(Value))^2)),
            Min = min(Value),
            Max = max(Value))

kable(df3)
```


# 3. Initial Exploration & Data Preprocessing

## 3.1 Time series

```{r}
#Time series
make_ts <- function(country,balance) {
df_balance <- df1 %>% filter(BALANCE == balance, Country==country) %>% 
  arrange(TIME) %>% drop_na(Value)
ts(df_balance$Value, start = c(2010, 1), frequency = 12)
}

ts_can_comb <- make_ts("Canada","COMBFUEL")
ts_can_nuc <- make_ts("Canada","NUCLEAR")
ts_can_hyd <- make_ts("Canada","HYDRO")
ts_usa_comb <- make_ts("United States","COMBFUEL")
ts_usa_nuc <- make_ts("United States","NUCLEAR")
ts_usa_hyd <- make_ts("United States","HYDRO")

#Fit linear trend
Tm_cc = as.vector(time(ts_can_comb))
out_cc = lm( ts_can_comb ~ Tm_cc )

Tm_cn = as.vector(time(ts_can_nuc))
out_cn = lm( ts_can_nuc ~ Tm_cn )

Tm_ch = as.vector(time(ts_can_hyd))
out_ch = lm( ts_can_hyd ~ Tm_ch )

Tm_uc = as.vector(time(ts_usa_comb))
out_uc = lm( ts_usa_comb ~ Tm_uc )

Tm_un = as.vector(time(ts_usa_nuc))
out_un = lm( ts_usa_nuc ~ Tm_un )

Tm_uh = as.vector(time(ts_usa_hyd))
out_uh = lm( ts_usa_hyd ~ Tm_uh )

#plot time series and find the trend
par(mfrow=c(3,2),mar=c(4,4,2,1),oma=c(0,0,3,0))
plot(ts_can_comb, ylab="Net Electricity Supply",main = "Canada COMBFUEL")
lines( Tm_cc, out_cc$fitted, type = "l", col = 2)
plot(ts_usa_comb, ylab="Net Electricity Supply",main = "USA COMBFUEL")
lines( Tm_uc, out_uc$fitted, type = "l", col = 2)
plot(ts_can_nuc, ylab="Net Electricity Supply",main = "Canada NUCLEAR")
lines( Tm_cn, out_cn$fitted, type = "l", col = 2)
plot(ts_usa_nuc, ylab="Net Electricity Supply",main = "USA NUCLEAR")
lines( Tm_un, out_un$fitted, type = "l", col = 2)
plot(ts_can_hyd, ylab="Net Electricity Supply",main = "Canada HYDRO")
lines( Tm_ch, out_ch$fitted, type = "l", col = 2)
plot(ts_usa_hyd, ylab="Net Electricity Supply",main = "USA HYDRO")
lines( Tm_uh, out_uh$fitted, type = "l", col = 2)
mtext("Graph A1: Time Series for Each Country and Balance", side=3, outer=TRUE, line=1.5, cex=1.2)
```

```{r}
plot_df <- df1 %>% mutate(time = ym(TIME) ) %>% arrange(Country, time)
ggplot(plot_df,
       aes(x = time, y = Value, colour = BALANCE)) +
  geom_line(linewidth = 0.73) +
  facet_wrap(~ Country, ncol = 1, scales = "free_y") +
  labs(title = "Graph A2: Energy mix time series by country",
       x = "Time", y = "Net electricity supply",
       colour = "Energy source") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  theme_minimal()
```
```{r}
ggplot(plot_df,
       aes(x = time, y = Value, fill = BALANCE)) +
  geom_area(position = "stack",alpha = 0.73) +
  facet_wrap(~ Country, ncol = 1, scales = "free_y") +
  labs(title = "Graph A3: Stacked electricity supply by energy source in Canada and the U.S.",
       x = "Time", y = "Net electricity supply",
       fill = "Energy source") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")
```
## 3.2 Trends and Seasonal Patterns in Net Electricity Supply by Energy Source
```{r}
#Decomposition (check trend and seasonality)

dcmp_cc=decompose(ts_can_comb)
dcmp_cn=decompose(ts_can_nuc)
dcmp_ch=decompose(ts_can_hyd)
dcmp_uc=decompose(ts_usa_comb)
dcmp_un=decompose(ts_usa_nuc)
dcmp_uh=decompose(ts_usa_hyd)

Can_Comb_trend = dcmp_cc$trend
USA_Comb_trend = dcmp_uc$trend
Can_NUC_trend = dcmp_cn$trend
USA_NUC_trend = dcmp_un$trend
Can_HYD_trend = dcmp_ch$trend
USA_HYD_trend = dcmp_uh$trend

Can_Comb_seasonal = dcmp_cc$seasonal
USA_Comb_seasonal = dcmp_uc$seasonal
Can_NUC_seasonal = dcmp_cn$seasonal
USA_NUC_seasonal = dcmp_un$seasonal
Can_HYD_seasonal = dcmp_ch$seasonal
USA_HYD_seasonal = dcmp_uh$seasonal


```

```{r}
library(zoo)   
ts_to_df <- function(ts_obj, comp_vec, country, balance, component){
  tibble(
    time_num = time(ts_obj),              
    value    = as.numeric(comp_vec),      
    Country  = country,
    BALANCE  = balance,
    component = component
  ) %>%
    drop_na(value) %>%                    
    mutate(time = as.Date(as.yearmon(time_num)))
}

trend_df <- bind_rows(
  ts_to_df(ts_can_comb, Can_Comb_trend, "Canada", "COMBFUEL", "Trend"),
  ts_to_df(ts_can_nuc,  Can_NUC_trend,  "Canada", "NUCLEAR", "Trend"),
  ts_to_df(ts_can_hyd,  Can_HYD_trend,  "Canada", "HYDRO", "Trend"),
  ts_to_df(ts_usa_comb, USA_Comb_trend, "United States", "COMBFUEL", "Trend"),
  ts_to_df(ts_usa_nuc,  USA_NUC_trend,  "United States", "NUCLEAR",  "Trend"),
  ts_to_df(ts_usa_hyd,  USA_HYD_trend,  "United States", "HYDRO", "Trend")
)

season_df <- bind_rows(
  ts_to_df(ts_can_comb, Can_Comb_seasonal, "Canada", "COMBFUEL", "Seasonal"),
  ts_to_df(ts_can_nuc,  Can_NUC_seasonal,  "Canada", "NUCLEAR", "Seasonal"),
  ts_to_df(ts_can_hyd,  Can_HYD_seasonal,  "Canada", "HYDRO", "Seasonal"),
  ts_to_df(ts_usa_comb, USA_Comb_seasonal, "United States", "COMBFUEL", "Seasonal"),
  ts_to_df(ts_usa_nuc,  USA_NUC_seasonal,  "United States", "NUCLEAR", "Seasonal"),
  ts_to_df(ts_usa_hyd,  USA_HYD_seasonal,  "United States", "HYDRO", "Seasonal")
)



ggplot(trend_df,
       aes(x = time, y = value, colour = BALANCE)) +
  geom_line(linewidth = 0.73) +
  facet_wrap(~ Country, ncol = 1, scales = "free_y") +
  labs(title  = "Graph A4: Long-term changes in the energy mix over time 
       (trend components)",
       x      = "Time",
       y      = "Trend in net electricity supply",
       colour = "Energy source") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  theme_minimal()



ggplot(season_df,
       aes(x = time, y = value, colour = BALANCE)) +
  geom_line(linewidth = 0.73) +
  facet_wrap(~ Country, ncol = 1, scales = "free_y") +
  labs(title  = "Graph A5: Seasonal variation in the energy mix over time 
       (seasonal components)",
       x      = "Time",
       y      = "Seasonal component",
       colour = "Energy source") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  theme_minimal()

```

*Note: In Graph A6 & A7, y-axis scale differs across panels*
```{r,out.width="60%",fig.align='center'}
combfuel = Can_Comb_seasonal; nuclear = Can_NUC_seasonal; hydro = Can_HYD_seasonal
plot(cbind(combfuel,nuclear,hydro),main="Graph A6: Seasonal component of Canada energy sources",
     cex.axis = 0.8,   
     cex.lab = 0.8)
combfuel = USA_Comb_seasonal; nuclear = USA_NUC_seasonal; hydro = USA_HYD_seasonal
plot(cbind(combfuel,nuclear,hydro),main="Graph A7: Seasonal component of United States energy sources ",
     cex.axis = 0.8,   
     cex.lab = 0.8)

```



## 3.3 Transformations and Differencing
```{r,eval=FALSE}
# prove variance is constant
plot(cbind(ts_can_nuc,diff(ts_can_nuc)),main = "")
```

\begin{center}
\textbf{Graph A8: Original and differenced nuclear supply series (Canada)}
\end{center}
```{r,echo=FALSE,out.width="60%",fig.align='center'}
# prove variance is constant
plot(cbind(ts_can_nuc,diff(ts_can_nuc)),main = "")

```
```{r,eval=F}
plot(cbind(ts_can_nuc,log(ts_can_nuc)),main = "")
```

\pagebreak

\begin{center}
\textbf{Graph A9: Original and transformed nuclear supply series (Canada)}
\end{center}
```{r,echo=F,out.width="60%",fig.align='center'}
plot(cbind(ts_can_nuc,log(ts_can_nuc)),main = "")
```

### 3.3.1 unit roots test
```{r}

library(tseries)
adf.test(ts_can_comb)
adf.test(ts_usa_comb)
adf.test(ts_can_nuc)
adf.test(ts_usa_nuc)
adf.test(ts_can_hyd)
adf.test(ts_usa_hyd)
```
### 3.3.2 Canada~Nuclear Unit Root Test After Differencing
```{r}
adf.test(diff(ts_can_nuc))
```

## 3.4 Autocorrelation, Partial Autocorrelation, and Cross-correlation

### 3.4.1 ACF plot and underlying seasonality 
```{r}
#ACF
par(mfrow=c(3,2),mar=c(2,2,3,1),oma=c(2,2,2,0))
acf(as.numeric(ts_can_comb),lag.max = 80,main="Canada COMBFUEL",ylim = c(-0.5, 1))
acf(as.numeric(ts_usa_comb),lag.max = 80,main="USA COMBFUEL",ylim = c(-0.5, 1))
acf(as.numeric(ts_can_nuc),lag.max = 80,main="Canada NUCLEAR",ylim = c(-0.5, 1))
acf(as.numeric(ts_usa_nuc),lag.max = 80,main="USA NUCLEAR",ylim = c(-0.5, 1))
acf(as.numeric(ts_can_hyd),lag.max = 80,main="Canada HYDRO",ylim = c(-0.5, 1))
acf(as.numeric(ts_usa_hyd),lag.max = 80,main="USA HYDRO",ylim = c(-0.5, 1))
mtext("Graph A10: ACF for Each Country and Energy Sources", side=3, outer=TRUE, line=0.9, cex=1)
mtext("Lag", side = 1, outer = TRUE, line = 0.9, cex = 1, font = 2) 
mtext("ACF", side = 2, outer = TRUE, line = 0.9, cex = 1, font = 2)  
```

### 3.4.2 PACF plot 
```{r}
#pacf
par(mfrow=c(3,2),mar=c(2,2,3,1),oma=c(2,2,2,0))
pacf(as.numeric(ts_can_comb),lag.max = 40,main="Canada COMBFUEL",ylim = c(-0.5, 1))
pacf(as.numeric(ts_usa_comb),lag.max = 40,main="USA COMBFUEL",ylim = c(-0.5, 1))
pacf(as.numeric(ts_can_nuc),lag.max = 40,main="Canada NUCLEAR",ylim = c(-0.5, 1))
pacf(as.numeric(ts_usa_nuc),lag.max = 40,main="USA NUCLEAR",ylim = c(-0.5, 1))
pacf(as.numeric(ts_can_hyd),lag.max = 40,main="Canada HYDRO",ylim = c(-0.5, 1))
pacf(as.numeric(ts_usa_hyd),lag.max = 40,main="USA HYDRO",ylim = c(-0.5, 1))
mtext("Graph A11: PACF for Each Country and Energy Sources", side=3, outer=TRUE, line=0.9, cex=1)
mtext("Lag", side = 1, outer = TRUE, line = 0.9, cex = 1, font = 2)
mtext("PACF", side = 2, outer = TRUE, line = 0.9, cex = 1, font = 2)
```

### 3.4.3 cross correlation

```{r}
par(mfrow=c(3,2),mar=c(2,2,3,1),oma=c(2,2,2,0))
ccf(as.numeric(ts_can_comb),as.numeric(ts_can_nuc),lag.max = 24, main = "Can Comb & Nuc")
ccf(as.numeric(ts_usa_comb),as.numeric(ts_usa_nuc),lag.max = 24, main = "USA Comb & Nuc")
ccf(as.numeric(ts_can_comb),as.numeric(ts_can_hyd),lag.max = 24, main = "Can Comb & Hdy")
ccf(as.numeric(ts_usa_comb),as.numeric(ts_usa_hyd),lag.max = 24, main = "USA Comb & Hdy")
ccf(as.numeric(ts_can_nuc),as.numeric(ts_can_hyd),lag.max = 24, main = "Can Nuc & Hdy")
ccf(as.numeric(ts_usa_nuc),as.numeric(ts_usa_hyd),lag.max = 24, main = "USA Nuc & Hdy")
mtext("Graph A12: CCF for Energy Sources in Same Country", side=3, outer=TRUE, line=0.9, cex=1)
mtext("Lag", side = 1, outer = TRUE, line = 0.9, cex = 1, font = 2)
mtext("CCF", side = 2, outer = TRUE, line = 0.9, cex = 1, font = 2)
```
```{r}
par(mfrow=c(2,2),mar=c(2,2,3,1),oma=c(2,2,2,0))
ccf(as.numeric(ts_can_comb),as.numeric(ts_usa_comb),lag.max = 40,main="Comb Can & USA")
ccf(as.numeric(ts_can_nuc),as.numeric(ts_usa_nuc),lag.max = 40,main="Nuc Can & USA")
ccf(as.numeric(ts_can_hyd),as.numeric(ts_usa_hyd),lag.max = 40,main="Hdy Can & USA")
mtext("Graph A13: CCF for Country in Same Energy Source", side=3, outer=TRUE, line=0.9, cex=1)
mtext("Lag", side = 1, outer = TRUE, line = 0.9, cex = 1, font = 2)
mtext("CCF", side = 2, outer = TRUE, line = 0.9, cex = 1, font = 2)
```

### 3.4.4 cross correlation with seasonal differencing
```{r}
# seasonal differencing
sd_can_c = seasadj(stl(ts_can_comb, s.window="periodic"))
sd_can_n = seasadj(stl(ts_can_nuc, s.window="periodic"))
sd_can_h = seasadj(stl(ts_can_hyd, s.window="periodic"))
sd_usa_c = seasadj(stl(ts_usa_comb, s.window="periodic"))
sd_usa_n = seasadj(stl(ts_usa_nuc, s.window="periodic"))
sd_usa_h = seasadj(stl(ts_usa_hyd, s.window="periodic"))

par(mfrow=c(3,2),mar=c(2,2,3,1),oma=c(2,2,2,0))
ccf(as.numeric(sd_can_c),as.numeric(sd_can_n),lag.max = 24, main = "Can Comb & Nuc")
ccf(as.numeric(sd_usa_c),as.numeric(sd_usa_n),lag.max = 24, main = "USA Comb & Nuc")
ccf(as.numeric(sd_can_c),as.numeric(sd_can_h),lag.max = 24, main = "Can Comb & Hdy")
ccf(as.numeric(sd_usa_c),as.numeric(sd_usa_h),lag.max = 24, main = "USA Comb & Hdy")
ccf(as.numeric(sd_can_n),as.numeric(sd_usa_h),lag.max = 24, main = "Can Nuc & Hdy")
ccf(as.numeric(sd_usa_n),as.numeric(sd_usa_h),lag.max = 24, main = "USA Nuc & Hdy")
mtext("Graph A14: Seasonal Differencing CCF for Energy Sources in Same Country", side=3, outer=TRUE, line=0.9, cex=1)
mtext("Lag", side = 1, outer = TRUE, line = 0.9, cex = 1, font = 2)
mtext("CCF", side = 2, outer = TRUE, line = 0.9, cex = 1, font = 2)

```
```{r}
par(mfrow=c(2,2),mar=c(2,2,3,1),oma=c(2,2,2,0))
ccf(as.numeric(sd_can_c),as.numeric(sd_usa_c),lag.max = 40,main="Comb Can & USA")
ccf(as.numeric(sd_can_n),as.numeric(sd_usa_n),lag.max = 40,main="Nuc Can & USA")
ccf(as.numeric(sd_can_h),as.numeric(sd_usa_h),lag.max = 40,main="Hdy Can & USA")
mtext("Graph A15: Seasonal Differencing CCF for Country in Same Energy Source", side=3, outer=TRUE, line=0.9, cex=1)
mtext("Lag", side = 1, outer = TRUE, line = 0.9, cex = 1, font = 2)
mtext("CCF", side = 2, outer = TRUE, line = 0.9, cex = 1, font = 2)
```


# 4 Modeling

## 4.1 Implementation & Model Evaluation
### 4.1.1 Univariate SARIMA Models
### Parameter Tuning 
#### SARIMA(stepwise AIC model selection)
```{r}

ts_list <- list(
  "Can COMB" = ts_can_comb,
  "USA COMB" = ts_usa_comb,
  "Can NUC"  = ts_can_nuc,
  "USA NUC"  = ts_usa_nuc,
  "Can HYD"  = ts_can_hyd,
  "USA HYD"  = ts_usa_hyd
)
fit_one <- function(x){
  auto.arima(head(x,-12), # use training set
             seasonal = TRUE, D = 1, ic = "aic",
             stepwise = TRUE, approximation = FALSE,
             allowmean = TRUE, allowdrift = TRUE)
}

fits <- map(ts_list, fit_one)



```

Use stepwise AIC criteria to determine parameters estimation and model order:
```{r,eval=FALSE}
summ_tbl <- imap(fits, function(m, nm){
  ord <- arimaorder(m)  # c(p,d,q,P,D,Q,s)

  order_chr    <- sprintf("(%d,%d,%d)",       ord[1], ord[2], ord[3])
  seasonal_chr <- sprintf("(%d,%d,%d)[%d]",   ord[4], ord[5], ord[6], ord[7])

  n_train <- length(m$x)                      
  aic_val <- AIC(m)


  tibble(
    Series = nm,
    Order = order_chr,
    Seasonal = seasonal_chr,
    AIC = round(aic_val, 2),
  )
})

model_table <- bind_rows(summ_tbl)

kable(
  model_table
)

```

\begin{center}
\textbf{Table A2: Stepwise AIC Selected SARIMA Model Orders}
\end{center}
```{r,echo=FALSE}

summ_tbl <- imap(fits, function(m, nm){
  ord <- arimaorder(m)  # c(p,d,q,P,D,Q,s)

  order_chr    <- sprintf("(%d,%d,%d)",       ord[1], ord[2], ord[3])
  seasonal_chr <- sprintf("(%d,%d,%d)[%d]",   ord[4], ord[5], ord[6], ord[7])

  n_train <- length(m$x)                      
  aic_val <- AIC(m)


  tibble(
    Series = nm,
    Order = order_chr,
    Seasonal = seasonal_chr,
    AIC = round(aic_val, 2),
  )
})

model_table <- bind_rows(summ_tbl)

kable(
  model_table
)

```


### 4.1.2 SARIMA Diagnostics
```{r,eval=FALSE}
library(astsa)
sarima(head(ts_can_comb,-12),1,0,1,0,1,1,S = 12)
mtext("SARIMA Diagnostics for Canada COMBFUEL",
      side = 3, line = 3.4, cex = 0.8, font = 2)
sarima(head(ts_usa_comb,-12),1,0,0,2,1,0,S = 12)
mtext("SARIMA Diagnostics for United States COMBFUEL",
      side = 3, line = 3.4, cex = 0.8, font = 2)
sarima(head(ts_can_nuc,-12),0,1,2,0,1,1,S = 12)
mtext("SARIMA Diagnostics for Canada NUCLEAR",
      side = 3, line = 3.4, cex = 0.8, font = 2)
sarima(head(ts_usa_nuc,-12),2,0,1,0,1,2,S = 12)
mtext("SARIMA Diagnostics for United States NUCLEAR",
      side = 3, line = 3.4, cex = 0.8, font = 2)
sarima(head(ts_can_hyd,-12),3,0,0,2,1,1,S = 12)
mtext("SARIMA Diagnostics for Canada HYDRO",
      side = 3, line = 3.4, cex = 0.8, font = 2)
sarima(head(ts_usa_hyd,-12),1,0,0,2,1,0,S = 12)
mtext("SARIMA Diagnostics for United States HYDRO",
      side = 3, line = 3.4, cex = 0.8, font = 2)

```

\begin{center}
\textbf{Graph A16: SARIMA Diagnostics plot of model given by stepwise AIC}
\end{center}
```{r,echo=FALSE,out.width="80%",fig.align='center'}
library(astsa)
sarima(head(ts_can_comb,-12),1,0,1,0,1,1,S = 12)
mtext("SARIMA Diagnostics for Canada COMBFUEL",
      side = 3, line = 3.4, cex = 0.8, font = 2)
sarima(head(ts_usa_comb,-12),1,0,0,2,1,0,S = 12)
mtext("SARIMA Diagnostics for United States COMBFUEL",
      side = 3, line = 3.4, cex = 0.8, font = 2)
sarima(head(ts_can_nuc,-12),0,1,2,0,1,1,S = 12)
mtext("SARIMA Diagnostics for Canada NUCLEAR",
      side = 3, line = 3.4, cex = 0.8, font = 2)
sarima(head(ts_usa_nuc,-12),2,0,1,0,1,2,S = 12)
mtext("SARIMA Diagnostics for United States NUCLEAR",
      side = 3, line = 3.4, cex = 0.8, font = 2)
sarima(head(ts_can_hyd,-12),3,0,0,2,1,1,S = 12)
mtext("SARIMA Diagnostics for Canada HYDRO",
      side = 3, line = 3.4, cex = 0.8, font = 2)
sarima(head(ts_usa_hyd,-12),1,0,0,2,1,0,S = 12)
mtext("SARIMA Diagnostics for United States HYDRO",
      side = 3, line = 3.4, cex = 0.8, font = 2)

```

### 4.1.3 SARIMA Model improvement
```{r,eval=FALSE}
# After modifying order which are base on diagnostics 

sarima(head(ts_usa_comb,-12),1,0,0,2,1,1,S = 12) #add sma terms
mtext("Improved SARIMA Diagnostics for United States COMBFUEL",
      side = 3, line = 3.4, cex = 0.8, font = 2)
sarima(head(ts_usa_hyd,-12),1,0,0,2,1,1,S = 12) #add sma terms
mtext("Improved SARIMA Diagnostics for United States HYDRO",
      side = 3, line = 3.4, cex = 0.8, font = 2)
```

\begin{center}
\textbf{Graph A17: Diagnostics plot of model after improvement}
\end{center}
```{r,echo=FALSE,out.width="60%"}
# After modifying order which are base on diagnostics 

sarima(head(ts_usa_comb,-12),1,0,0,2,1,1,S = 12) #add sma terms
mtext("Improved SARIMA Diagnostics for United States COMBFUEL",
      side = 3, line = 3.4, cex = 0.8, font = 2)
sarima(head(ts_usa_hyd,-12),1,0,0,2,1,1,S = 12) #add sma terms
mtext("Improved SARIMA Diagnostics for United States HYDRO",
      side = 3, line = 3.4, cex = 0.8, font = 2)
```

```{r}
fit_cc = Arima(head(ts_can_comb,-12),order = c(1,0,1), seasonal = list(order = c(0,1,1),
               period = 12),include.drift = T)
fit_uc = Arima(head(ts_usa_comb,-12),order = c(1,0,0), seasonal = list(order = c(2,1,1),
               period = 12),include.drift = T)
fit_cn = Arima(head(ts_can_nuc,-12),order = c(0,1,2), seasonal = list(order = c(0,1,1),
               period = 12),include.drift = T)
fit_un = Arima(head(ts_usa_nuc,-12),order = c(2,0,1), seasonal = list(order = c(0,1,2),
               period = 12),include.drift = T)
fit_ch = Arima(head(ts_can_hyd,-12),order = c(3,0,0), seasonal = list(order = c(2,1,1),
               period = 12),include.drift = T)
fit_uh = Arima(head(ts_usa_hyd,-12),order = c(1,0,0), seasonal = list(order = c(2,1,1),
               period = 12),include.drift = T)
```

### 4.1.4 Multivariate VAR Models

### Parameter Tuning

### VAR model with all series
```{r}
Y = cbind(C_Comb = ts_can_comb,
          C_Nuc  = ts_can_nuc,
          C_Hyd  = ts_can_hyd,
          U_Comb = ts_usa_comb,
          U_Nuc  = ts_usa_nuc,
          U_Hyd  = ts_usa_hyd)
VARselect(Y, lag.max = 5, type = "both",season = 12)
VAR(Y, type = "both", ic = "AIC", lag.max =5, season = 12)
```

AIC is big, we choose 2 different way to consider multivariate relationship

### Within-country VAR (three-variable VAR models for each country)
We adopt a three-variable VAR model within each country because the CCF results 
indicate strong contemporaneous interdependence among COMBFUEL, HYDRO, and NUCLEAR 
supplies. Modeling them jointly improves forecast performance and allows us to 
quantify how changes in one source help predict the others.

so in this case, we use traning set to fit model
```{r}
library(vars)
Y_can = cbind(Comb = ts_can_comb,
              Nuc  = ts_can_nuc,
              Hyd  = ts_can_hyd)

Y_usa = cbind(Comb = ts_usa_comb,
              Nuc  = ts_usa_nuc,
              Hyd  = ts_usa_hyd)

Y_can_train <- cbind(Comb = head(ts_can_comb,-12),
               Nuc  = head(ts_can_nuc,-12),
               Hyd  = head(ts_can_hyd,-12))

Y_usa_train <- cbind(Comb = head(ts_usa_comb,-12),
               Nuc  = head(ts_usa_nuc,-12),
               Hyd  = head(ts_usa_hyd,-12))

# Decide order p
VARselect(Y_can_train, lag.max = 5, type = "both",season = 12)$selection
(fit_var_can = VAR(Y_can_train, type = "both", ic = "AIC", lag.max = 5,season = 12))

VARselect(Y_usa_train, lag.max = 5, type = "both",season = 12)$selection
(fit_var_usa = VAR(Y_usa_train, type = "both", ic = "AIC", lag.max = 5,season = 12))

```
So we choose VAR(3) in Canada, and VAR(1) in United States.

### Cross-country bivariate VAR for the same source
```{r}
Y_comb = cbind(Can = ts_can_comb,
               USA = ts_usa_comb)
Y_nuc = cbind(Can = ts_can_nuc,
              USA = ts_usa_nuc)
Y_hyd = cbind(Can = ts_can_hyd,
              USA = ts_usa_hyd)

VARselect(Y_comb, lag.max = 5, type = "both",season = 12)$selection
(fit_var_comb = VAR(Y_comb, type = "both", ic = "AIC", lag.max = 5,season = 12))
VARselect(Y_nuc, lag.max = 5, type = "both",season = 12)$selection
(fit_var_nuc = VAR(Y_nuc, type = "both", ic = "AIC", lag.max = 5,season = 12))
VARselect(Y_hyd, lag.max = 5, type = "both",season = 12)$selection
(fit_var_hyd = VAR(Y_hyd, type = "both", ic = "AIC", lag.max = 5,season = 12))
```

### 4.1.5 Seasonal Differenced VAR model 
#### Seasonal Differenced Within-country VAR
```{r}
Y_can_sd = cbind(Comb = sd_can_c,
                 Nuc  = sd_can_n,
                 Hyd  = sd_can_h)

Y_usa_sd = cbind(Comb = sd_usa_c,
                 Nuc  = sd_usa_n,
                 Hyd  = sd_usa_h)
VARselect(Y_can_sd, lag.max = 5, type = "both")$selection
(fit_var_can_sd = VAR(Y_can_sd, type = "both", ic = "AIC", lag.max = 5))

VARselect(Y_usa_sd, lag.max = 5, type = "both")$selection
(fit_var_usa_sd = VAR(Y_usa_sd, type = "both", ic = "AIC", lag.max = 5))
```
After seasonal differencing, we choose VAR(3) for Canada, and VAR(2) for United States.

#### Seasonal Differenced Cross-country bivariate VAR
```{r}
Y_comb_sd = cbind(Can = sd_can_c,
                  USA = sd_usa_c)
Y_nuc_sd = cbind(Can = sd_can_n,
                 USA = sd_usa_n)
Y_hyd_sd = cbind(Can = sd_can_h,
                 USA = sd_usa_h)

VARselect(Y_comb_sd, lag.max = 5, type = "both")$selection
(fit_var_comb_sd = VAR(Y_comb_sd, type = "both", ic = "AIC", lag.max = 5))
VARselect(Y_nuc_sd, lag.max = 5, type = "both")$selection
(fit_var_nuc_sd = VAR(Y_nuc_sd, type = "both", ic = "AIC", lag.max = 5))
VARselect(Y_hyd_sd, lag.max = 5, type = "both")$selection
(fit_var_hyd_sd = VAR(Y_hyd_sd, type = "both", ic = "AIC", lag.max = 5))
```
After seasonal differencing, we choose VAR(2) for Combfuel, Nuclear, and VAR(4) for Hydro.

### 4.1.6 VAR Diagnostics
### Within-country VAR

```{r,eval=FALSE}
# get residuals 
res_can = resid(fit_var_can) %>% ts( start = c(2010, 1), freq = 12)
res_usa = resid(fit_var_usa) %>% ts( start = c(2010, 1), freq = 12)
# Residual plot
plot(res_can, main = "3-variable VAR model residuals plots in Canada")
plot(res_usa, main = "3-variable VAR model residuals plots in United States")
```

\begin{center}
\textbf{Graph A18: Residuals CCF of 3-variable VAR}
\end{center}
```{r,echo=FALSE,out.width="55%"}
# get residuals 
res_can = resid(fit_var_can) %>% ts( start = c(2010, 1), freq = 12)
res_usa = resid(fit_var_usa) %>% ts( start = c(2010, 1), freq = 12)
# Residual plot
plot(res_can, main = "3-variable VAR model residuals plots in Canada")
plot(res_usa, main = "3-variable VAR model residuals plots in United States")
```

\pagebreak

```{r,eval=FALSE}
# Residual ACF/CCF plots
acf(res_can)
```

\begin{center}
\textbf{Graph A19: Residuals CCF of 3-variable VAR in Canada}
\end{center}
```{r,echo=FALSE}
# Residual ACF/CCF plots
acf(res_can)

```

\pagebreak

```{r,eval=FALSE}
acf(res_usa)
```

\begin{center}
\textbf{Graph A20: Residuals CCF of 3-variable VAR in United States}
\end{center}
```{r,echo=FALSE}
acf(res_usa)
```

\pagebreak

```{r,eval=FALSE}
# CAN Normal QQ-plots
par(mfrow = c(2,2))
qqnorm(res_can[,"Comb"]); qqline(res_can[,"Comb"])
qqnorm(res_can[,"Nuc"]); qqline(res_can[,"Nuc"])
qqnorm(res_can[,"Hyd"]); qqline(res_can[,"Hyd"])

```

\begin{center}
\textbf{Graph A21: Q-Q plot for 3-variable VAR model in Canada}
\end{center}
```{r,echo=FALSE}
# CAN Normal QQ-plots
par(mfrow = c(2,2))
qqnorm(res_can[,"Comb"]); qqline(res_can[,"Comb"])
qqnorm(res_can[,"Nuc"]); qqline(res_can[,"Nuc"])
qqnorm(res_can[,"Hyd"]); qqline(res_can[,"Hyd"])

```

\pagebreak

```{r,eval=FALSE}
# USA Normal QQ-plots
par(mfrow = c(2,2))
qqnorm(res_usa[,"Comb"]); qqline(res_usa[,"Comb"])
qqnorm(res_usa[,"Nuc"]); qqline(res_usa[,"Nuc"])
qqnorm(res_usa[,"Hyd"]); qqline(res_usa[,"Hyd"])
```

\begin{center}
\textbf{Graph A22: Q-Q plot for 3-variable VAR model in United States}
\end{center}
```{r,echo=FALSE}
# USA Normal QQ-plots
par(mfrow = c(2,2))
qqnorm(res_usa[,"Comb"]); qqline(res_usa[,"Comb"])
qqnorm(res_usa[,"Nuc"]); qqline(res_usa[,"Nuc"])
qqnorm(res_usa[,"Hyd"]); qqline(res_usa[,"Hyd"])
```

\pagebreak

### Cross-country VAR
```{r,eval=FALSE}
# get residuals 
res_comb = resid(fit_var_comb) %>% ts( start = c(2010, 1), freq = 12)
res_nuc = resid(fit_var_nuc) %>% ts( start = c(2010, 1), freq = 12)
res_hyd = resid(fit_var_hyd) %>% ts( start = c(2010, 1), freq = 12)
# Residual plot
plot(res_comb, main = "2-variable VAR model residuals plots in Combfuel")
plot(res_nuc, main = "2-variable VAR model residuals plots in Nuclear")
plot(res_hyd, main = "2-variable VAR model residuals plots in Hydro")
```

\begin{center}
\textbf{Graph A23: 2-variable VAR model residuals plots}
\end{center}
```{r,echo=FALSE,out.width="55%"}
# get residuals 
res_comb = resid(fit_var_comb) %>% ts( start = c(2010, 1), freq = 12)
res_nuc = resid(fit_var_nuc) %>% ts( start = c(2010, 1), freq = 12)
res_hyd = resid(fit_var_hyd) %>% ts( start = c(2010, 1), freq = 12)
# Residual plot
plot(res_comb, main = "2-variable VAR model residuals plots in Combfuel")
plot(res_nuc, main = "2-variable VAR model residuals plots in Nuclear")
plot(res_hyd, main = "2-variable VAR model residuals plots in Hydro")
```

\pagebreak

```{r,eval=FALSE}
# Residual ACF/CCF plots
acf(res_comb)
```

\begin{center}
\textbf{Graph A24: Residuals CCF for 2-variable VAR model in Combfuel}
\end{center}
```{r,echo=FALSE}
# Residual ACF/CCF plots
acf(res_comb)
```

\pagebreak

```{r,eval=FALSE}
acf(res_nuc)
```

\begin{center}
\textbf{Graph A25: Residuals CCF for 2-variable VAR model in Nuclear}
\end{center}
```{r,echo=FALSE}
acf(res_nuc)
```

\pagebreak

```{r,eval=FALSE}
acf(res_hyd)
```

\begin{center}
\textbf{Graph A26: Residuals CCF for 2-variable VAR model in Hydro}
\end{center}
```{r,echo=FALSE}
acf(res_hyd)
```

\pagebreak

```{r,eval=FALSE}
par(mfrow = c(1,2))
qqnorm(res_comb[,"Can"]); qqline(res_comb[,"Can"])
qqnorm(res_comb[,"USA"]); qqline(res_comb[,"USA"])
```

\begin{center}
\textbf{Graph A27: Q-Q plot for 2-variable VAR model in Combfuel}
\end{center}
```{r,echo=FALSE}
par(mfrow = c(1,2))
qqnorm(res_comb[,"Can"]); qqline(res_comb[,"Can"])
qqnorm(res_comb[,"USA"]); qqline(res_comb[,"USA"])

```

\pagebreak

```{r,eval=FALSE}
par(mfrow = c(1,2))
qqnorm(res_nuc[,"Can"]); qqline(res_nuc[,"Can"])
qqnorm(res_nuc[,"USA"]); qqline(res_nuc[,"USA"])
```

\begin{center}
\textbf{Graph A28: Q-Q plot for 2-variable VAR model in Nuclear}
\end{center}
```{r,echo=FALSE}
par(mfrow = c(1,2))
qqnorm(res_nuc[,"Can"]); qqline(res_nuc[,"Can"])
qqnorm(res_nuc[,"USA"]); qqline(res_nuc[,"USA"])
```

\pagebreak

```{r,eval=FALSE}
par(mfrow = c(1,2))
qqnorm(res_hyd[,"Can"]); qqline(res_hyd[,"Can"])
qqnorm(res_hyd[,"USA"]); qqline(res_hyd[,"USA"])
```

\begin{center}
\textbf{Graph A29: Q-Q plot for 2-variable VAR model in Hydro}
\end{center}
```{r,echo=FALSE}
par(mfrow = c(1,2))
qqnorm(res_hyd[,"Can"]); qqline(res_hyd[,"Can"])
qqnorm(res_hyd[,"USA"]); qqline(res_hyd[,"USA"])
```

### 4.1.7 Seasonal Differenced VAR Diagnostics
#### Within-country VAR

```{r,eval=FALSE}
# get residuals 
res_can = resid(fit_var_can_sd) %>% ts( start = c(2010, 1), freq = 12)
res_usa = resid(fit_var_usa_sd) %>% ts( start = c(2010, 1), freq = 12)
# Residual plot
plot(res_can, main = "3-variable VAR model residuals plots in Canada")
plot(res_usa, main = "3-variable VAR model residuals plots in United States")
```

\begin{center}
\textbf{Graph A30: Seasonal Differenced Residuals CCF of 3-variable VAR}
\end{center}
```{r,echo=FALSE,out.width="55%"}
# get residuals 
res_can_sd = resid(fit_var_can_sd) %>% ts( start = c(2010, 1), freq = 12)
res_usa_sd = resid(fit_var_usa_sd) %>% ts( start = c(2010, 1), freq = 12)
# Residual plot
plot(res_can_sd, main = "3-variable VAR model residuals plots in Canada")
plot(res_usa_sd, main = "3-variable VAR model residuals plots in United States")
```

\pagebreak

```{r,eval=FALSE}
# Residual ACF/CCF plots
acf(res_can_sd)
```

\begin{center}
\textbf{Graph A31: Seasonal Differenced Residuals CCF of 3-variable VAR in Canada}
\end{center}
```{r,echo=FALSE}
# Residual ACF/CCF plots
acf(res_can_sd)

```

\pagebreak

```{r,eval=FALSE}
acf(res_usa_sd)
```

\begin{center}
\textbf{Graph A32: Seasonal Differenced Residuals CCF of 3-variable VAR in United States}
\end{center}
```{r,echo=FALSE}
acf(res_usa_sd)
```

\pagebreak

```{r,eval=FALSE}
# CAN Normal QQ-plots
par(mfrow = c(2,2))
qqnorm(res_can_sd[,"Comb"]); qqline(res_can_sd[,"Comb"])
qqnorm(res_can_sd[,"Nuc"]); qqline(res_can_sd[,"Nuc"])
qqnorm(res_can_sd[,"Hyd"]); qqline(res_can_sd[,"Hyd"])

```

\begin{center}
\textbf{Graph A33: Seasonal Differenced Q-Q plot for 3-variable VAR model in Canada}
\end{center}
```{r,echo=FALSE}
# CAN Normal QQ-plots
par(mfrow = c(2,2))
qqnorm(res_can_sd[,"Comb"]); qqline(res_can_sd[,"Comb"])
qqnorm(res_can_sd[,"Nuc"]); qqline(res_can_sd[,"Nuc"])
qqnorm(res_can_sd[,"Hyd"]); qqline(res_can_sd[,"Hyd"])

```

\pagebreak

```{r,eval=FALSE}
# USA Normal QQ-plots
par(mfrow = c(2,2))
qqnorm(res_usa_sd[,"Comb"]); qqline(res_usa_sd[,"Comb"])
qqnorm(res_usa_sd[,"Nuc"]); qqline(res_usa_sd[,"Nuc"])
qqnorm(res_usa_sd[,"Hyd"]); qqline(res_usa_sd[,"Hyd"])
```

\begin{center}
\textbf{Graph A34: Seasonal Differenced Q-Q plot for 3-variable VAR model in United States}
\end{center}
```{r,echo=FALSE}
# USA Normal QQ-plots
par(mfrow = c(2,2))
qqnorm(res_usa_sd[,"Comb"]); qqline(res_usa_sd[,"Comb"])
qqnorm(res_usa_sd[,"Nuc"]); qqline(res_usa_sd[,"Nuc"])
qqnorm(res_usa_sd[,"Hyd"]); qqline(res_usa_sd[,"Hyd"])
```

\pagebreak

#### Cross-country VAR
```{r,eval=FALSE}
# get residuals 
res_comb_sd = resid(fit_var_comb_sd) %>% ts( start = c(2010, 1), freq = 12)
res_nuc_sd = resid(fit_var_nuc_sd) %>% ts( start = c(2010, 1), freq = 12)
res_hyd_sd = resid(fit_var_hyd_sd) %>% ts( start = c(2010, 1), freq = 12)
# Residual plot
plot(res_comb_sd, main = "2-variable VAR model residuals plots in Combfuel")
plot(res_nuc_sd, main = "2-variable VAR model residuals plots in Nuclear")
plot(res_hyd_sd, main = "2-variable VAR model residuals plots in Hydro")
```

\begin{center}
\textbf{Graph A35: Seasonal Differenced 2-variable VAR model residuals plots}
\end{center}
```{r,echo=FALSE,out.width="55%"}
# get residuals 
res_comb_sd = resid(fit_var_comb_sd) %>% ts( start = c(2010, 1), freq = 12)
res_nuc_sd = resid(fit_var_nuc_sd) %>% ts( start = c(2010, 1), freq = 12)
res_hyd_sd = resid(fit_var_hyd_sd) %>% ts( start = c(2010, 1), freq = 12)
# Residual plot
plot(res_comb_sd, main = "2-variable VAR model residuals plots in Combfuel")
plot(res_nuc_sd, main = "2-variable VAR model residuals plots in Nuclear")
plot(res_hyd_sd, main = "2-variable VAR model residuals plots in Hydro")
```

\pagebreak

```{r,eval=FALSE}
# Residual ACF/CCF plots
acf(res_comb_sd)
```

\begin{center}
\textbf{Graph A36: Seasonal Differenced Residuals CCF for 2-variable VAR model in Combfuel}
\end{center}
```{r,echo=FALSE}
# Residual ACF/CCF plots
acf(res_comb_sd)
```

\pagebreak

```{r,eval=FALSE}
acf(res_nuc_sd)
```

\begin{center}
\textbf{Graph A37: Seasonal Differenced Residuals CCF for 2-variable VAR model in Nuclear}
\end{center}
```{r,echo=FALSE}
acf(res_nuc_sd)
```

\pagebreak

```{r,eval=FALSE}
acf(res_hyd_sd)
```

\begin{center}
\textbf{Graph A38: Seasonal Differenced Residuals CCF for 2-variable VAR model in Hydro}
\end{center}
```{r,echo=FALSE}
acf(res_hyd_sd)
```

\pagebreak

```{r,eval=FALSE}
par(mfrow = c(1,2))
qqnorm(res_comb_sd[,"Can"]); qqline(res_comb_sd[,"Can"])
qqnorm(res_comb_sd[,"USA"]); qqline(res_comb_sd[,"USA"])
```

\begin{center}
\textbf{Graph A39: Seasonal Differenced Q-Q plot for 2-variable VAR model in Combfuel}
\end{center}
```{r,echo=FALSE}
par(mfrow = c(1,2))
qqnorm(res_comb_sd[,"Can"]); qqline(res_comb_sd[,"Can"])
qqnorm(res_comb_sd[,"USA"]); qqline(res_comb_sd[,"USA"])

```

\pagebreak

```{r,eval=FALSE}
par(mfrow = c(1,2))
qqnorm(res_nuc_sd[,"Can"]); qqline(res_nuc_sd[,"Can"])
qqnorm(res_nuc_sd[,"USA"]); qqline(res_nuc_sd[,"USA"])
```

\begin{center}
\textbf{Graph A40: Seasonal Differenced Q-Q plot for 2-variable VAR model in Nuclear}
\end{center}
```{r,echo=FALSE}
par(mfrow = c(1,2))
qqnorm(res_nuc_sd[,"Can"]); qqline(res_nuc_sd[,"Can"])
qqnorm(res_nuc_sd[,"USA"]); qqline(res_nuc_sd[,"USA"])
```

\pagebreak

```{r,eval=FALSE}
par(mfrow = c(1,2))
qqnorm(res_hyd_sd[,"Can"]); qqline(res_hyd_sd[,"Can"])
qqnorm(res_hyd_sd[,"USA"]); qqline(res_hyd_sd[,"USA"])
```

\begin{center}
\textbf{Graph A41: Seasonal Differenced Q-Q plot for 2-variable VAR model in Hydro}
\end{center}
```{r,echo=FALSE}
par(mfrow = c(1,2))
qqnorm(res_hyd_sd[,"Can"]); qqline(res_hyd_sd[,"Can"])
qqnorm(res_hyd_sd[,"USA"]); qqline(res_hyd_sd[,"USA"])
```


# 5. Model Evaluation
## Validation approach(forecast accuracy)
## 5.1 Graph validation forecast
### SARIMA model validation forecast
```{r,eval=FALSE}
forecast( fit_cc, 12 ) %>% plot(main = "Canada SARIMA Validation Forecast for Comb")
lines( tail( ts_can_comb, 12 ), lwd=2, col =2)
legend( "topleft", legend = c("actual", "forecast"), col = c(2,4), lwd = 2)
forecast( fit_cn, 12 ) %>% plot(main = "Canada SARIMA Validation Forecast for Nuc")
lines( tail( ts_can_nuc, 12 ), lwd=2, col =2)
legend( "topleft", legend = c(main = "actual", "forecast"), col = c(2,4), lwd = 2)
forecast( fit_ch, 12 ) %>% plot(main = "Canada SARIMA Validation Forecast for Hyd")
lines( tail( ts_can_hyd, 12 ), lwd=2, col =2)
legend( "topleft", legend = c("actual", "forecast"), col = c(2,4), lwd = 2)
forecast( fit_uc, 12 ) %>% plot(main = "Unite State SARIMA Validation Forecast for Comb")
lines( tail( ts_usa_comb, 12 ), lwd=2, col =2)
legend( "topleft", legend = c("actual", "forecast"), col = c(2,4), lwd = 2)
forecast( fit_un, 12 ) %>% plot(main = "Unite State SARIMA Validation Forecast for Nuc")
lines( tail( ts_usa_nuc, 12 ), lwd=2, col =2)
legend( "topleft", legend = c("actual", "forecast"), col = c(2,4), lwd = 2)
forecast( fit_uh, 12 ) %>% plot(main = "Unite State SARIMA Validation Forecast for Hyd")
lines( tail( ts_usa_hyd, 12 ), lwd=2, col =2)
legend( "topleft", legend = c("actual", "forecast"), col = c(2,4), lwd = 2)
```

\begin{center}
\textbf{Graph A42: SARIMA Validation Forecast}
\end{center}
```{r,echo=FALSE}
forecast( fit_cc, 12 ) %>% plot(main = "Canada SARIMA Validation Forecast for Comb")
lines( tail( ts_can_comb, 12 ), lwd=2, col =2)
legend( "topleft", legend = c("actual", "forecast"), col = c(2,4), lwd = 2)
forecast( fit_cn, 12 ) %>% plot(main = "Canada SARIMA Validation Forecast for Nuc")
lines( tail( ts_can_nuc, 12 ), lwd=2, col =2)
legend( "topleft", legend = c(main = "actual", "forecast"), col = c(2,4), lwd = 2)
forecast( fit_ch, 12 ) %>% plot(main = "Canada SARIMA Validation Forecast for Hyd")
lines( tail( ts_can_hyd, 12 ), lwd=2, col =2)
legend( "topleft", legend = c("actual", "forecast"), col = c(2,4), lwd = 2)
forecast( fit_uc, 12 ) %>% plot(main = "Unite State SARIMA Validation Forecast for Comb")
lines( tail( ts_usa_comb, 12 ), lwd=2, col =2)
legend( "topleft", legend = c("actual", "forecast"), col = c(2,4), lwd = 2)
forecast( fit_un, 12 ) %>% plot(main = "Unite State SARIMA Validation Forecast for Nuc")
lines( tail( ts_usa_nuc, 12 ), lwd=2, col =2)
legend( "topleft", legend = c("actual", "forecast"), col = c(2,4), lwd = 2)
forecast( fit_uh, 12 ) %>% plot(main = "Unite State SARIMA Validation Forecast for Hyd")
lines( tail( ts_usa_hyd, 12 ), lwd=2, col =2)
legend( "topleft", legend = c("actual", "forecast"), col = c(2,4), lwd = 2)
```

### VAR model validation forecast`
```{r,eval=FALSE}
fc <- forecast(fit_var_can, 12)
fc1 <- forecast(fit_var_usa, 12)
series_names <- colnames(Y_can)
series_names1 <- colnames(Y_usa)
for(i in 1:3){
  plot(fc$forecast[[i]], main = paste("Canada VAR Validation Forecast for", series_names[i]))
  lines(tail(Y_can[, i], 12), col = "red", lwd = 2)
  legend("topleft", legend = c("actual", "forecast"),
         col = c("red", "blue"), lwd = 2)
}

for(i in 1:3){
  plot(fc1$forecast[[i]], main = paste("Unite State VAR Validation Forecast for", series_names1[i]))
  lines(tail(Y_usa[, i], 12), col = "red", lwd = 2)
  legend("topleft", legend = c("actual", "forecast"),
         col = c("red", "blue"), lwd = 2)
}

```

\pagebreak

\begin{center}
\textbf{Graph A43: VAR Validation Forecast}
\end{center}
```{r,echo=FALSE}
fc <- forecast(fit_var_can, 12)
fc1 <- forecast(fit_var_usa, 12)
series_names <- colnames(Y_can)
series_names1 <- colnames(Y_usa)
for(i in 1:3){
  plot(fc$forecast[[i]], main = paste("Canada VAR Validation Forecast for", series_names[i]))
  lines(tail(Y_can[, i], 12), col = "red", lwd = 2)
  legend("topleft", legend = c("actual", "forecast"),
         col = c("red", "blue"), lwd = 2)
}

for(i in 1:3){
  plot(fc1$forecast[[i]], main = paste("Unite State VAR Validation Forecast for", series_names1[i]))
  lines(tail(Y_usa[, i], 12), col = "red", lwd = 2)
  legend("topleft", legend = c("actual", "forecast"),
         col = c("red", "blue"), lwd = 2)
}

```



## 5.2 Key performance metrics(RMSE, MAE)

### SARIMA model RMSE & MAE
```{r,eval=FALSE}
calc_accuracy <- function(ts_data, fit_obj, h = 12) {
  valid <- tail(ts_data, h)            # validation 
  fc <- forecast(fit_obj, h = h)       
  pred <- fc$mean                      
  
  errors <- valid - pred
  RMSE <- sqrt(mean(errors^2, na.rm = TRUE))
  MAE  <- mean(abs(errors), na.rm = TRUE)
  
  return(c(RMSE = RMSE, MAE = MAE))
}



fit_list <- list(
  "Can COMB" = fit_cc,
  "USA COMB" = fit_uc,
  "Can NUC"  = fit_cn,
  "USA NUC"  = fit_un,
  "Can HYD"  = fit_ch,
  "USA HYD"  = fit_uh
)

results <- map2_df(ts_list, fit_list, calc_accuracy, .id = "Series")

kable(results)
```

\begin{center}
\textbf{Table A3: Each time series SARIMA RMSE and MAE}
\end{center}
```{r,echo=FALSE}
calc_accuracy <- function(ts_data, fit_obj, h = 12) {
  valid <- tail(ts_data, h)            # validation 
  fc <- forecast(fit_obj, h = h)       
  pred <- fc$mean                      
  
  errors <- valid - pred
  RMSE <- sqrt(mean(errors^2, na.rm = TRUE))
  MAE  <- mean(abs(errors), na.rm = TRUE)
  
  return(c(RMSE = RMSE, MAE = MAE))
}



fit_list <- list(
  "Can COMB" = fit_cc,
  "USA COMB" = fit_uc,
  "Can NUC"  = fit_cn,
  "USA NUC"  = fit_un,
  "Can HYD"  = fit_ch,
  "USA HYD"  = fit_uh
)

results <- map2_df(ts_list, fit_list, calc_accuracy, .id = "Series")

kable(results)
```

Although USA COMB and USA HYD show relatively large RMSE values in absolute scale, 
this is mainly due to their substantially higher data magnitudes. Their forecast trajectories 
closely follow the true observations, and relative accuracy metrics (MAPE) remain acceptable. 
Therefore, no further model refinement is required, avoiding unnecessary risk of overfitting.

### VAR model RMSE & MAE
```{r,eval=FALSE}
calc_accuracy_var <- function(ts_data, fit_var, var_index, h = 12) {
  
  valid <- tail(ts_data, h)        
  fc    <- forecast(fit_var, h = h)
  pred  <- fc$forecast[[var_index]]$mean   
  
  errors <- valid - pred
  RMSE   <- sqrt(mean(errors^2, na.rm = TRUE))
  MAE    <- mean(abs(errors), na.rm = TRUE)
  
  c(RMSE = RMSE, MAE = MAE)
}

fit_var_list <- list(
  fit_var_can,
  fit_var_usa,
  fit_var_can,
  fit_var_usa,
  fit_var_can,
  fit_var_usa
)

var_index <- c(1, 1, 2, 2, 3, 3)
series_names <- names(ts_list)

results_var <- pmap_dfr(
  list(ts_data   = ts_list,
       fit_var   = fit_var_list,
       var_index = var_index,
       Series    = series_names),
  ~{
    acc <- calc_accuracy_var(..1, ..2, ..3, h = 12)
    tibble(Series = ..4,
           RMSE   = acc["RMSE"],
           MAE    = acc["MAE"])
  }
)

kable(results_var)

```

\begin{center}
\textbf{Table A4: Each time series VAR RMSE and MAE}
\end{center}
```{r,echo=FALSE}
calc_accuracy_var <- function(ts_data, fit_var, var_index, h = 12) {
  
  valid <- tail(ts_data, h)        
  fc    <- forecast(fit_var, h = h)
  pred  <- fc$forecast[[var_index]]$mean   
  
  errors <- valid - pred
  RMSE   <- sqrt(mean(errors^2, na.rm = TRUE))
  MAE    <- mean(abs(errors), na.rm = TRUE)
  
  c(RMSE = RMSE, MAE = MAE)
}

fit_var_list <- list(
  fit_var_can,
  fit_var_usa,
  fit_var_can,
  fit_var_usa,
  fit_var_can,
  fit_var_usa
)

var_index <- c(1, 1, 2, 2, 3, 3)
series_names <- names(ts_list)

results_var <- pmap_dfr(
  list(ts_data   = ts_list,
       fit_var   = fit_var_list,
       var_index = var_index,
       Series    = series_names),
  ~{
    acc <- calc_accuracy_var(..1, ..2, ..3, h = 12)
    tibble(Series = ..4,
           RMSE   = acc["RMSE"],
           MAE    = acc["MAE"])
  }
)

kable(results_var)

```


# 6. Final model for Forecasting 
Final Training (Refit best model) use all data(training + validation)
```{r,eval=FALSE}
final_cc = Arima(ts_can_comb,order = c(1,0,1), seasonal = list(order = c(1,1,1),
               period = 12),include.drift = T)
final_uc = VAR(Y_usa, type = "both", ic = "AIC", p = 1, season = 12)$varresult[[1]]
final_cn = Arima(ts_can_nuc,order = c(0,1,2), seasonal = list(order = c(0,1,1),
               period = 12),include.drift = T)
final_un = Arima(ts_usa_nuc,order = c(2,0,1), seasonal = list(order = c(0,1,2),
               period = 12),include.drift = T)
final_ch = Arima(ts_can_hyd,order = c(3,0,0), seasonal = list(order = c(2,1,1),
               period = 12),include.drift = T)
final_uh = Arima(ts_usa_hyd,order = c(1,0,0), seasonal = list(order = c(2,1,1),
               period = 12),include.drift = T)

model_list <- list(
  "CAN COMB" = final_cc,
  "USA COMB" = final_uc,
  "CAN NUC"  = final_cn,
  "USA NUC"  = final_un,
  "CAN HYD" = final_ch,
  "USA HYD" = final_uh
)
summ_tbl1 <- imap(model_list, function(m, nm){
  if ("Arima" %in% class(m)) {

    ord <- arimaorder(m)
    order_chr    <- sprintf("SARIMA(%d,%d,%d)(%d,%d,%d)[%d]", ord[1], ord[2], ord[3],
                            ord[4], ord[5], ord[6], ord[7])
    aic_val  <- AIC(m)

    

  
  } else if ("lm" %in% class(m)) {

    
    order_chr    <- "VAR(1,seasonal = 12)"
    aic_val      <- AIC(m)
    

  } else {
    stop(sprintf("Unsupported model class for %s: %s",
                 nm, paste(class(m), collapse = ", ")))
  }

  tibble(
    Series = nm,
    Order = order_chr,
    AIC = round(aic_val, 2)
  )
})

final_table <- bind_rows(summ_tbl1)

kable(
  final_table
  )
```

\begin{center}
\textbf{Table A5: Final forecasting model summary}
\end{center}
```{r,warning=FALSE,echo=FALSE}
final_cc = Arima(ts_can_comb,order = c(1,0,1), seasonal = list(order = c(1,1,1),
               period = 12),include.drift = T)
final_uc = VAR(Y_usa, type = "both", ic = "AIC", p = 1, season = 12)$varresult[[1]]
final_cn = Arima(ts_can_nuc,order = c(0,1,2), seasonal = list(order = c(0,1,1),
               period = 12),include.drift = T)
final_un = Arima(ts_usa_nuc,order = c(2,0,1), seasonal = list(order = c(0,1,2),
               period = 12),include.drift = T)
final_ch = Arima(ts_can_hyd,order = c(3,0,0), seasonal = list(order = c(2,1,1),
               period = 12),include.drift = T)
final_uh = Arima(ts_usa_hyd,order = c(1,0,0), seasonal = list(order = c(2,1,1),
               period = 12),include.drift = T)

model_list <- list(
  "CAN COMB" = final_cc,
  "USA COMB" = final_uc,
  "CAN NUC"  = final_cn,
  "USA NUC"  = final_un,
  "CAN HYD" = final_ch,
  "USA HYD" = final_uh
)
summ_tbl1 <- imap(model_list, function(m, nm){
  if ("Arima" %in% class(m)) {

    ord <- arimaorder(m)
    order_chr    <- sprintf("SARIMA(%d,%d,%d)(%d,%d,%d)[%d]", ord[1], ord[2], ord[3],
                            ord[4], ord[5], ord[6], ord[7])
    aic_val  <- AIC(m)

    

  
  } else if ("lm" %in% class(m)) {

    
    order_chr    <- "VAR(1,seasonal = 12)"
    aic_val      <- AIC(m)
    

  } else {
    stop(sprintf("Unsupported model class for %s: %s",
                 nm, paste(class(m), collapse = ", ")))
  }

  tibble(
    Series = nm,
    Order = order_chr,
    AIC = round(aic_val, 2)
  )
})

final_table <- bind_rows(summ_tbl1)

kable(
  final_table
  )
```
Note: the VAR model has different AIC criteria with SARIMA model since they have 
different constant (k).

## Coefficients for Final forecasting model
```{r,eval=FALSE}
collapse_coef <- function(fit){
  cf <- coef(fit)
  paste0(names(cf), " = ", round(cf, 3), collapse = ", ")
}


coef_cc <- collapse_coef(final_cc)
coef_uc <- collapse_coef(final_uc)
coef_cn <- collapse_coef(final_cn)
coef_un <- collapse_coef(final_un)
coef_ch <- collapse_coef(final_ch)
coef_uh <- collapse_coef(final_uh)

# Build final table
coef_table <- data.frame(
  Series = c(
    "Canada Comb",
    "USA Comb",
    "Canada Nuclear",
    "USA Nuclear",
    "Canada Hydro",
    "USA Hydro"
  ),
  Coefficients = c(
    coef_cc,
    coef_uc,
    coef_cn,
    coef_un,
    coef_ch,
    coef_uh
  ),
  stringsAsFactors = FALSE
)

kable(coef_table,escape = FALSE)
```

\begin{center}
\textbf{Table A6: Final forecasting model coefficients summary}
\end{center}
```{r,echo=FALSE}
collapse_coef <- function(fit){
  cf <- coef(fit)
  parts <- paste0(names(cf), " = ", round(cf, 3))
  paste(parts, collapse=", ")
}


coef_cc <- collapse_coef(final_cc)
coef_uc <- collapse_coef(final_uc)
coef_cn <- collapse_coef(final_cn)
coef_un <- collapse_coef(final_un)
coef_ch <- collapse_coef(final_ch)
coef_uh <- collapse_coef(final_uh)

# Build final table
coef_table <- data.frame(
  Series = c(
    "Canada Comb",
    "USA Comb",
    "Canada Nuclear",
    "USA Nuclear",
    "Canada Hydro",
    "USA Hydro"
  ),
  Coefficients = c(
    coef_cc,
    coef_uc,
    coef_cn,
    coef_un,
    coef_ch,
    coef_uh
  ),
  stringsAsFactors = FALSE
)

kable(coef_table,escape = FALSE)
```


# 6.1 Forecastion Findings(Q2)
```{r,eval=FALSE}
final_fc = VAR(Y_usa, type = "both", ic = "AIC", p = 1, season = 12)
fc_uc = forecast(final_fc, 24)
forecast( final_cc, 24 ) %>% plot(main = "Canada SARIMA Forecast for Combfuel")
plot(fc_uc$forecast[[1]], main = paste("Unite State VAR Forecast for Combfuel", series_names1[1]))
forecast( final_cn, 24 ) %>% plot(main = "Canada SARIMA Forecast for Nuclear")
forecast( final_un, 24 ) %>% plot(main = "Unite State SARIMA Forecast For Nuclear")
forecast( final_ch, 24 ) %>% plot(main = "Canada SARIMA Forecast for Hydro")
forecast( final_uh, 24 ) %>% plot(main = "Unite State SARIMA Forecast for Hydro")
```

\begin{center}
\textbf{Graph A44: Each Series 2-years Forecat Plots}
\end{center}
```{r,echo=FALSE}
final_fc = VAR(Y_usa, type = "both", ic = "AIC", p = 1, season = 12)
fc_uc = forecast(final_fc, 24)
forecast( final_cc, 24 ) %>% plot(main = "Canada SARIMA Forecast for Combfuel")
plot(fc_uc$forecast[[1]], main = paste("Unite State VAR Forecast for Combfuel", series_names1[1]))
forecast( final_cn, 24 ) %>% plot(main = "Canada SARIMA Forecast for Nuclear")
forecast( final_un, 24 ) %>% plot(main = "Unite State SARIMA Forecast For Nuclear")
forecast( final_ch, 24 ) %>% plot(main = "Canada SARIMA Forecast for Hydro")
forecast( final_uh, 24 ) %>% plot(main = "Unite State SARIMA Forecast for Hydro")
```



# 7. Final model for Interdependence Analysis
```{r}
VARselect(Y_can, lag.max = 5, type = "both",season = 12)$selection
VARselect(Y_usa, lag.max = 5, type = "both",season = 12)$selection
```

```{r,eval=FALSE}
fit_var_can = VAR(Y_can, type = "both", ic = "AIC", p = 1,season = 12)

fit_var_usa = VAR(Y_usa, type = "both", ic = "AIC", p = 1, season = 12)

aic_can <- round(AIC(fit_var_can), 2)
aic_usa <- round(AIC(fit_var_usa), 2)

var_summary <- tibble(
  Country = c("Canada", "USA"),
  Model = c("VAR(1)", "VAR(1)"),
  Seasonal = c("Seasonal = 12", "Seasonal = 12"),
  Parameters = c(
    "const, drift, Comb.l1, Nuc.l1, Hyd.l1",
    "const, drift, Comb.l1, Nuc.l1, Hyd.l1"
  ),
  AIC = c(aic_can, aic_usa)
)

kable(var_summary
) 


```

\begin{center}
\textbf{Table A7: Selected 3-variable VAR models for Canada and USA}
\end{center}
```{r,echo=FALSE}

fit_var_can = VAR(Y_can, type = "both", ic = "AIC", p = 1,season = 12)

fit_var_usa = VAR(Y_usa, type = "both", ic = "AIC", p = 1, season = 12)

aic_can <- round(AIC(fit_var_can), 2)
aic_usa <- round(AIC(fit_var_usa), 2)

var_summary <- tibble(
  Country = c("Canada", "USA"),
  Model = c("VAR(1)", "VAR(1)"),
  Seasonal = c("Seasonal = 12", "Seasonal = 12"),
  Parameters = c(
    "const, drift, Comb.l1, Nuc.l1, Hyd.l1",
    "const, drift, Comb.l1, Nuc.l1, Hyd.l1"
  ),
  AIC = c(aic_can, aic_usa)
)

kable(var_summary
) 



```

## coefficients in these model

```{r,eval=FALSE}
collapse_var_eq <- function(fit, eq){
  cf_mat <- coef(fit)[[eq]]          
  est    <- cf_mat[, 1]              

  parts <- paste0(rownames(cf_mat), " = ", round(est, 2))
  paste(parts, collapse = ", ")
}
coef_can_comb <- collapse_var_eq(fit_var_can, "Comb")
coef_usa_comb <- collapse_var_eq(fit_var_usa, "Comb")

coef_can_nuc  <- collapse_var_eq(fit_var_can, "Nuc")
coef_usa_nuc  <- collapse_var_eq(fit_var_usa, "Nuc")

coef_can_hyd  <- collapse_var_eq(fit_var_can, "Hyd")
coef_usa_hyd  <- collapse_var_eq(fit_var_usa, "Hyd")

coef_table_var <- tibble(
  Series = c(
    "Can Comb",
    "USA Comb",
    "Can Nuclear",
    "USA Nuclear",
    "Can Hydro",
    "USA Hydro"
  ),
  Coefficients = c(
    coef_can_comb,
    coef_usa_comb,
    coef_can_nuc,
    coef_usa_nuc,
    coef_can_hyd,
    coef_usa_hyd
  )
)

kable(coef_table_var, escape = FALSE)
```


\begin{center}
\textbf{Table A8: Selected 3-variable VAR models coefficients for Canada and USA}
\end{center}
```{r,echo=FALSE}
collapse_var_eq <- function(fit, eq){
  cf_mat <- coef(fit)[[eq]]          
  est    <- cf_mat[, 1]              

  parts <- paste0(rownames(cf_mat), " = ", round(est, 2))
  paste(parts, collapse = ", ")
}
coef_can_comb <- collapse_var_eq(fit_var_can, "Comb")
coef_usa_comb <- collapse_var_eq(fit_var_usa, "Comb")

coef_can_nuc  <- collapse_var_eq(fit_var_can, "Nuc")
coef_usa_nuc  <- collapse_var_eq(fit_var_usa, "Nuc")

coef_can_hyd  <- collapse_var_eq(fit_var_can, "Hyd")
coef_usa_hyd  <- collapse_var_eq(fit_var_usa, "Hyd")

coef_table_var <- tibble(
  Series = c(
    "Canada Comb",
    "USA Comb",
    "Canada Nuclear",
    "USA Nuclear",
    "Canada Hydro",
    "USA Hydro"
  ),
  Coefficients = c(
    coef_can_comb,
    coef_usa_comb,
    coef_can_nuc,
    coef_usa_nuc,
    coef_can_hyd,
    coef_usa_hyd
  )
)

kable(coef_table_var, escape = FALSE)
```


```{r,eval=FALSE}
fit_var_comb = VAR(Y_comb, type = "both", ic = "AIC", p = 2, season = 12)
fit_var_nuc = VAR(Y_nuc, type = "both", ic = "AIC", p = 2,season = 12)
fit_var_hyd = VAR(Y_hyd, type = "both", ic = "AIC", p = 4, season = 12)

var_summary_cc <- tibble(
  Energy = c("Combfuel", "Nuclear", "Hydro"),
  Model = c("VAR(2)", "VAR(2)", "VAR(4)"),
  Seasonal = c("Seasonal = 12", "Seasonal = 12", "Seasonal = 12"),
  Parameters = c(
    "Lag1-2 of all variables + Const + Trend",
    "Lag1-2 of all variables + Const + Trend",
    "Lag1-4 of all variables + Const + Trend"
  ),
  AIC = c(
    AIC(fit_var_comb) %>% round(2),
    AIC(fit_var_nuc) %>% round(2),
    AIC(fit_var_hyd) %>% round(2)
  )
)

kable(var_summary_cc
) 
```

\begin{center}
\textbf{Table A9: Selected 2-variable VAR models for Combfuel, Nuclear, Hydro}
\end{center}
```{r,echo=FALSE}
fit_var_comb = VAR(Y_comb, type = "both", ic = "AIC", p = 2, season = 12)
fit_var_nuc = VAR(Y_nuc, type = "both", ic = "AIC", p = 2,season = 12)
fit_var_hyd = VAR(Y_hyd, type = "both", ic = "AIC", p = 4, season = 12)

var_summary_cc <- tibble(
  Energy = c("Combfuel", "Nuclear", "Hydro"),
  Model = c("VAR(2)", "VAR(2)", "VAR(4)"),
  Seasonal = c("Seasonal = 12", "Seasonal = 12", "Seasonal = 12"),
  Parameters = c(
    "Lag1-2 of all variables + Const + Trend",
    "Lag1-2 of all variables + Const + Trend",
    "Lag1-4 of all variables + Const + Trend"
  ),
  AIC = c(
    AIC(fit_var_comb) %>% round(2),
    AIC(fit_var_nuc) %>% round(2),
    AIC(fit_var_hyd) %>% round(2)
  )
)

kable(var_summary_cc
) 

```

## coefficients in these model

```{r,eval=FALSE}
coef_can_comb <- collapse_var_eq(fit_var_comb, "Can")
coef_usa_comb <- collapse_var_eq(fit_var_comb, "USA")

coef_can_nuc  <- collapse_var_eq(fit_var_nuc, "Can")
coef_usa_nuc  <- collapse_var_eq(fit_var_nuc, "USA")

coef_can_hyd  <- collapse_var_eq(fit_var_hyd, "Can")
coef_usa_hyd  <- collapse_var_eq(fit_var_hyd, "USA")

coef_table_var <- tibble(
  Series = c(
    "Canada Comb",
    "USA Comb",
    "Canada Nuclear",
    "USA Nuclear",
    "Canada Hydro",
    "USA Hydro"
  ),
  Coefficients = c(
    coef_can_comb,
    coef_usa_comb,
    coef_can_nuc,
    coef_usa_nuc,
    coef_can_hyd,
    coef_usa_hyd
  )
)

kable(coef_table_var, escape = FALSE)
```
\begin{center}
\textbf{Table A9: Selected 2-variable VAR models coefficients for Combfuel, Nuclear, Hydro}
\end{center}
```{r,echo=FALSE}
coef_can_comb <- collapse_var_eq(fit_var_comb, "Can")
coef_usa_comb <- collapse_var_eq(fit_var_comb, "USA")

coef_can_nuc  <- collapse_var_eq(fit_var_nuc, "Can")
coef_usa_nuc  <- collapse_var_eq(fit_var_nuc, "USA")

coef_can_hyd  <- collapse_var_eq(fit_var_hyd, "Can")
coef_usa_hyd  <- collapse_var_eq(fit_var_hyd, "USA")

coef_table_var <- tibble(
  Series = c(
    "Canada Comb",
    "USA Comb",
    "Canada Nuclear",
    "USA Nuclear",
    "Canada Hydro",
    "USA Hydro"
  ),
  Coefficients = c(
    coef_can_comb,
    coef_usa_comb,
    coef_can_nuc,
    coef_usa_nuc,
    coef_can_hyd,
    coef_usa_hyd
  )
)

kable(coef_table_var, escape = FALSE)
```

## Seasonal Differenced VAR model
```{r,eval=FALSE}
aic_can_sd <- round(AIC(fit_var_can_sd), 2)
aic_usa_sd <- round(AIC(fit_var_usa_sd), 2)

var_summary_sd <- tibble(
  Country = c("Canada", "USA"),
  Model = c("VAR(3)", "VAR(2)"),
  Parameters = c(
    "Lag1-2 of all variables + Const + Trend",
    "Lag1-3 of all variables + Const + Trend"
  ),
  AIC = c(aic_can_sd, aic_usa_sd)
)

kable(var_summary_sd) 
```

\begin{center}
\textbf{Table A10: Selected Seasonal Differenced 3-variable VAR models for Canada and USA}
\end{center}
```{r,echo=FALSE}
aic_can_sd <- round(AIC(fit_var_can_sd), 2)
aic_usa_sd <- round(AIC(fit_var_usa_sd), 2)

var_summary_sd <- tibble(
  Country = c("Canada", "USA"),
  Model = c("VAR(3)", "VAR(2)"),
  Parameters = c(
    "Lag1-2 of all variables + Const + Trend",
    "Lag1-3 of all variables + Const + Trend"
  ),
  AIC = c(aic_can_sd, aic_usa_sd)
)

kable(var_summary_sd) 
```

```{r,eval=FALSE}
aic_c_sd=round(AIC(fit_var_comb_sd),2)
aic_n_sd=round(AIC(fit_var_nuc_sd),2)
aic_h_sd=round(AIC(fit_var_nuc_sd),2)

var_summary_cc_sd <- tibble(
  Energy = c("Combfuel", "Nuclear", "Hydro"),
  Model = c("VAR(2)", "VAR(2)", "VAR(4)"),
  Parameters = c(
    "Lag1-2 of all variables + Const + Trend",
    "Lag1-2 of all variables + Const + Trend",
    "Lag1-4 of all variables + Const + Trend"
  ),
  AIC = c(aic_c_sd,aic_n_sd,aic_h_sd
  )
)

kable(var_summary_cc_sd) 
```

\begin{center}
\textbf{Table A11: Selected Seasonal Differenced 2-variable VAR models for Combfuel, Nuclear, Hydro}
\end{center}
```{r,echo=FALSE}

aic_c_sd=round(AIC(fit_var_comb_sd),2)
aic_n_sd=round(AIC(fit_var_nuc_sd),2)
aic_h_sd=round(AIC(fit_var_nuc_sd),2)

var_summary_cc_sd <- tibble(
  Energy = c("Combfuel", "Nuclear", "Hydro"),
  Model = c("VAR(2)", "VAR(2)", "VAR(1)"),
  Parameters = c(
    "Lag1-2 of all variables + Const + Trend",
    "Lag1-2 of all variables + Const + Trend",
    "Lag1 of all variables + Const + Trend"
  ),
  AIC = c(aic_c_sd,aic_n_sd,aic_h_sd
  )
)

kable(var_summary_cc_sd) 
```

### coefficients
```{r,eval=FALSE}
coef_can_comb_sd <- collapse_var_eq(fit_var_can_sd, "Comb")
coef_usa_comb_sd <- collapse_var_eq(fit_var_usa_sd, "Comb")

coef_can_nuc_sd  <- collapse_var_eq(fit_var_can_sd, "Nuc")
coef_usa_nuc_sd  <- collapse_var_eq(fit_var_usa_sd, "Nuc")

coef_can_hyd_sd  <- collapse_var_eq(fit_var_can_sd, "Hyd")
coef_usa_hyd_sd  <- collapse_var_eq(fit_var_usa_sd, "Hyd")

coef_table_var_sd <- tibble(
  Series = c(
    "Canada Comb",
    "USA Comb",
    "Canada Nuclear",
    "USA Nuclear",
    "Canada Hydro",
    "USA Hydro"
  ),
  Coefficients = c(
    coef_can_comb_sd,
    coef_usa_comb_sd,
    coef_can_nuc_sd,
    coef_usa_nuc_sd,
    coef_can_hyd_sd,
    coef_usa_hyd_sd
  )
)

kable(coef_table_var_sd, escape = FALSE)
```

\pagebreak

\begin{center}
\textbf{Table A12: Seasonal Differenced 3-variable VAR models coefficients for Canada and USA}
\end{center}
```{r,echo=FALSE}
coef_can_comb_sd <- collapse_var_eq(fit_var_can_sd, "Comb")
coef_usa_comb_sd <- collapse_var_eq(fit_var_usa_sd, "Comb")

coef_can_nuc_sd  <- collapse_var_eq(fit_var_can_sd, "Nuc")
coef_usa_nuc_sd  <- collapse_var_eq(fit_var_usa_sd, "Nuc")

coef_can_hyd_sd  <- collapse_var_eq(fit_var_can_sd, "Hyd")
coef_usa_hyd_sd  <- collapse_var_eq(fit_var_usa_sd, "Hyd")

coef_table_var_sd <- tibble(
  Series = c(
    "Canada Comb",
    "USA Comb",
    "Canada Nuclear",
    "USA Nuclear",
    "Canada Hydro",
    "USA Hydro"
  ),
  Coefficients = c(
    coef_can_comb_sd,
    coef_usa_comb_sd,
    coef_can_nuc_sd,
    coef_usa_nuc_sd,
    coef_can_hyd_sd,
    coef_usa_hyd_sd
  )
)

kable(coef_table_var_sd, escape = FALSE)
```

```{r,eval=FALSE}
coef_can_comb_sd <- collapse_var_eq(fit_var_comb_sd, "Can")
coef_usa_comb_sd <- collapse_var_eq(fit_var_comb_sd, "USA")

coef_can_nuc_sd  <- collapse_var_eq(fit_var_nuc_sd, "Can")
coef_usa_nuc_sd  <- collapse_var_eq(fit_var_nuc_sd, "USA")

coef_can_hyd_sd  <- collapse_var_eq(fit_var_hyd_sd, "Can")
coef_usa_hyd_sd  <- collapse_var_eq(fit_var_hyd_sd, "USA")

coef_table_var_sd <- tibble(
  Series = c(
    "Canada Comb",
    "USA Comb",
    "Canada Nuclear",
    "USA Nuclear",
    "Canada Hydro",
    "USA Hydro"
  ),
  Coefficients = c(
    coef_can_comb_sd,
    coef_usa_comb_sd,
    coef_can_nuc_sd,
    coef_usa_nuc_sd,
    coef_can_hyd_sd,
    coef_usa_hyd_sd
  )
)

kable(coef_table_var_sd, escape = FALSE)
```


\begin{center}
\textbf{Table A13:Seasonal Differenced 2-variable VAR models cofficients for Combfuel, Nuclear, Hydro}
\end{center}
```{r,echo=FALSE}
coef_can_comb_sd <- collapse_var_eq(fit_var_comb_sd, "Can")
coef_usa_comb_sd <- collapse_var_eq(fit_var_comb_sd, "USA")

coef_can_nuc_sd  <- collapse_var_eq(fit_var_nuc_sd, "Can")
coef_usa_nuc_sd  <- collapse_var_eq(fit_var_nuc_sd, "USA")

coef_can_hyd_sd  <- collapse_var_eq(fit_var_hyd_sd, "Can")
coef_usa_hyd_sd  <- collapse_var_eq(fit_var_hyd_sd, "USA")

coef_table_var_sd <- tibble(
  Series = c(
    "Canada Comb",
    "USA Comb",
    "Canada Nuclear",
    "USA Nuclear",
    "Canada Hydro",
    "USA Hydro"
  ),
  Coefficients = c(
    coef_can_comb_sd,
    coef_usa_comb_sd,
    coef_can_nuc_sd,
    coef_usa_nuc_sd,
    coef_can_hyd_sd,
    coef_usa_hyd_sd
  )
)

kable(coef_table_var_sd, escape = FALSE)
```


# 7.1 Correlation Findings(Q3)
## Granger and Instantaneous Causality
```{r}
# In Canada
causality(fit_var_can, cause = "Comb")
causality(fit_var_can, cause = "Nuc")
causality(fit_var_can, cause = "Hyd")
# In United State
causality(fit_var_usa, cause = "Comb")
causality(fit_var_usa, cause = "Nuc")
causality(fit_var_usa, cause = "Hyd")

causality(fit_var_comb, cause = "Can")  # Can -> USA 
causality(fit_var_comb, cause = "USA")  # USA -> Can 

causality(fit_var_nuc, cause = "Can")  # Can -> USA 
causality(fit_var_nuc, cause = "USA")  # USA -> Can 

causality(fit_var_hyd, cause = "Can")  # Can -> USA 
causality(fit_var_hyd, cause = "USA")  # USA -> Can 
```
Base on Granger Causality p value at 5% significant level, only in the energy Nuclear,
Canada has a Granger Cause to Unite State.

## Seasonal Differenced Granger Causality
```{r}
# In Canada
causality(fit_var_can_sd, cause = "Comb")
causality(fit_var_can_sd, cause = "Nuc")
causality(fit_var_can_sd, cause = "Hyd")
# In United State
causality(fit_var_usa_sd, cause = "Comb")
causality(fit_var_usa_sd, cause = "Nuc")
causality(fit_var_usa_sd, cause = "Hyd")

causality(fit_var_comb_sd, cause = "Can")  # Can -> USA 
causality(fit_var_comb_sd, cause = "USA")  # USA -> Can 

causality(fit_var_nuc_sd, cause = "Can")  # Can -> USA 
causality(fit_var_nuc_sd, cause = "USA")  # USA -> Can 

causality(fit_var_hyd_sd, cause = "Can")  # Can -> USA 
causality(fit_var_hyd_sd, cause = "USA")  # USA -> Can 
```


## Impluse Response Function 
```{r,eval=FALSE}
irf(fit_var_can, ortho = F) %>% plot()
```

\begin{center}
\textbf{Graph A45: Impluse Response from Energy Source in Canada}
\end{center}
```{r,echo=FALSE,out.width="60%"}
irf(fit_var_can, ortho = F) %>% plot()
```

```{r,eval=FALSE}
irf(fit_var_usa, ortho = F) %>% plot()
```

\begin{center}
\textbf{Graph A46: Impluse Response from Energy Source in United States}
\end{center}
```{r,echo=FALSE,out.width="60%"}
irf(fit_var_usa, ortho = F) %>% plot()
```

```{r,eval=FALSE}
irf(fit_var_comb, ortho = F) %>% plot()
```

\begin{center}
\textbf{Graph A47: Impluse Response from Country in Combfuel}
\end{center}
```{r,echo=FALSE,out.width="60%"}
irf(fit_var_comb, ortho = F) %>% plot()
```

```{r,eval=FALSE}
irf(fit_var_nuc, ortho = F) %>% plot()
```

\begin{center}
\textbf{Graph A48: Impluse Response from Country in Nuclear}
\end{center}
```{r,echo=FALSE,out.width="60%"}
irf(fit_var_nuc, ortho = F) %>% plot()
```

```{r,eval=FALSE}
irf(fit_var_hyd, ortho = F) %>% plot()
```

\begin{center}
\textbf{Graph A49: Impluse Response from Country in Hydro}
\end{center}
```{r,echo=FALSE,out.width="60%"}
irf(fit_var_hyd, ortho = F) %>% plot()
```

## Seasonal Differenced Impluse Response Function
```{r,eval=FALSE}
irf(fit_var_can_sd, ortho = F) %>% plot()
```

\begin{center}
\textbf{Graph A50: Impluse Response from Energy Source in Canada}
\end{center}
```{r,echo=FALSE,out.width="60%"}
irf(fit_var_can_sd, ortho = F) %>% plot()
```

```{r,eval=FALSE}
irf(fit_var_usa_sd, ortho = F) %>% plot()
```

\begin{center}
\textbf{Graph A51: Impluse Response from Energy Source in United States}
\end{center}
```{r,echo=FALSE,out.width="60%"}
irf(fit_var_usa_sd, ortho = F) %>% plot()
```

```{r,eval=FALSE}
irf(fit_var_comb_sd, ortho = F) %>% plot()
```

\begin{center}
\textbf{Graph A52: Impluse Response from Country in Combfuel}
\end{center}
```{r,echo=FALSE,out.width="60%"}
irf(fit_var_comb_sd, ortho = F) %>% plot()
```

```{r,eval=FALSE}
irf(fit_var_nuc_sd, ortho = F) %>% plot()
```

\begin{center}
\textbf{Graph A53: Impluse Response from Country in Nuclear}
\end{center}
```{r,echo=FALSE,out.width="60%"}
irf(fit_var_nuc_sd, ortho = F) %>% plot()
```

```{r,eval=FALSE}
irf(fit_var_hyd_sd, ortho = F) %>% plot()
```

\pagebreak

\begin{center}
\textbf{Graph A54: Impluse Response from Country in Hydro}
\end{center}
```{r,echo=FALSE,out.width="60%"}
irf(fit_var_hyd_sd, ortho = F) %>% plot()
```


